Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-10-13 ms.Rnw
some tidying/commenting code. haitipkg update needed

<<Setup, include=FALSE,echo=FALSE,results='hide'>>=
library(knitr)
library(pomp)
library(panelPomp)
library(spatPomp)
library(foreach)
library(doParallel)
library(doRNG)
library(haitipkg)
library(tidyverse)

RUN_LEVEL <- 1
rl_dir <- paste0("run_level_", RUN_LEVEL, "/")

# run_level_* for model 2 is not needed.
for (i in c(1, 3)) {
  if (!dir.exists(paste0("model", i, '/', rl_dir))) {
    dir.create(paste0("model", i, '/', rl_dir), recursive = TRUE)
  }
}

if (!dir.exists('model2')) dir.create("model2/")

@@

theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5)))

# 40 cores for doob, 8 for ito
# doob_cores <- 40
# gl_cores <- 36

cores <-  as.numeric(Sys.getenv('SLURM_NTASKS_PER_NODE', unset=NA))
if(is.na(cores)) cores <- detectCores()
registerDoParallel(cores)

dep_labeller <- as_labeller(
  c(
    'Artibonite' = 'Artibonite',
    'Sud_Est' = 'Sud-Est',
    'Sud.Est' = 'Sud-Est',
    'Nippes' = 'Nippes',
    'Nord_Est' = 'Nord-Est',
    'Nord.Est' = 'Nord-Est',
    'Ouest' = 'Ouest',
    'Centre' = 'Centre',
    'Nord' = 'Nord',
    'Sud' = 'Sud',
    'Nord_Ouest' = 'Nord-Ouest',
    'Nord.Ouest' = 'Nord-Ouest',
    'Grande_Anse' = 'Grand\'Anse',
    'Grand.Anse' = 'Grand\'Anse'
  )
)

plot_order <- c(
  'Artibonite',
  'Sud_Est',
  'Nippes',
  'Nord_Est',
  'Ouest',
  'Centre',
  'Nord',
  'Sud',
  'Nord_Ouest',
  'Grande_Anse'
)

@@

dep_plot_df <- haitiCholera %>%
  select(-report) %>%
  pivot_longer(
    data = .,
    cols = -date_saturday,
    values_to = 'cases',
    names_to = 'dep'
  ) %>%
  mutate(
    date = as.Date(date_saturday),
    dep = gsub("\\.", "_", dep)
  ) %>%
  mutate(
    dep = case_when(dep == "Grand_Anse" ~ "Grande_Anse", TRUE ~ dep)
  )

true_agg_cases <- dep_plot_df %>%
  tidyr::pivot_wider(
    data = .,
    id_cols = c(date),
    names_from = dep,
    values_from = cases,
    names_prefix = 'cases_'
  ) %>%
  mutate(
    ReportedAll = cases_Artibonite + cases_Centre +
      cases_Grande_Anse + cases_Nippes + cases_Nord +
      cases_Nord_Est + cases_Ouest + cases_Sud +
      cases_Sud_Est + cases_Nord_Ouest
  )
@

@@

set.seed(141837)
sample_beta_nu <- function(mu_beta = 5.8, mu_nu = 0.96, n_params = N_PARAMS) {
  # This function is used to sample starting values for beta and nu.
  #
  # params:
  #   mu_beta: Average of the marginal distribution for beta1. Value was fit visually.
  #   mu_nu: Average of the marginal distribution for nu. Value was fit visually.
  # returns:
  #   data.frame of size 300x2, values of beta1 and nu with truncated bivariate
  #   normal relation.

  # Determine a good value for the marginal standard deviations
  sd_beta <- (12.1-mu_beta) / qnorm(0.995)
  sd_nu <- (0.9-mu_nu) / qnorm(0.005)

  # Create covariance matrix for the bivariate-normal distribution
  cov_mat <- rbind(
    c(sd_beta^2, -0.7 * sd_beta * sd_nu),
    c(-0.7 * sd_beta * sd_nu, sd_nu^2)
  )

  # Sample from the bivariate normal distribution
  X <- MASS::mvrnorm(n = n_params, mu = c(mu_beta, mu_nu), Sigma = cov_mat)
  colnames(X) <- c("beta1", "nu")

  # Identify values that fell outside of range, so they can be resampled
  bad_beta <- which(X[, 'beta1'] > 12.1 | X[, 'beta1'] < 0.04)
  bad_nu <- which(X[, 'nu'] > 1 | X[, 'nu'] < 0.9)
  n_resamp <- length(unique(c(bad_beta, bad_nu)))

  # Keep resampling values until all values fall within the range.
  while (n_resamp > 0) {
    to_resamp <- unique(c(bad_beta, bad_nu))
    X[to_resamp, ] <- MASS::mvrnorm(n = n_resamp, mu = c(mu_beta, mu_nu), Sigma = cov_mat)

    bad_beta <- which(X[, 'beta1'] > 12.1 | X[, 'beta1'] < 0.04)
    bad_nu <- which(X[, 'nu'] > 1 | X[, 'nu'] < 0.9)
    n_resamp <- length(unique(c(bad_beta, bad_nu)))
  }

  # Return results as a data.frame
  as.data.frame(X)
}
# Run the function to get values of beta1 and nu
starting_params <- sample_beta_nu()

sample_rho <- function(beta, n_params = N_PARAMS) {
  # This function is used to get starting values for the parameter rho, based
  # on it's strong correlation with beta1 and nu. Here, we first sample beta1
  # and nu, and then sample rho based on a regression on beta1. Additionally,
  # it is noted that rho has a bi-modal distribution, with lower values of rho
  # corresponding with high probability to higher values of beta1.
  #
  # params:
  #   beta: previously sampled values of beta1 that will be used to get values
  #         of rho.
  # returns:
  #   vector of length 300 that have an appropriate relationship with the parameter
  #   beta1.

  rhos <- rep(0, N_PARAMS)

  # Get slope and intercept for the regression
  slope <- (0.4/(5.1 - 6.75))
  intercpt <- -6.75 * slope

  # Determine which mode should the value of rho belong to, based on regression
  case_when(
    beta > 6.75 ~ 0,  # Large beta --> group "0"
    # Small beta --> more likely to be group "1", with increasing probability as beta gets smaller
    beta <= 6.75 & beta > (0.95 - intercpt) / slope ~ slope * beta + intercpt,
    TRUE ~ 0.95
  ) -> probs

  # Sample the groups that rho belongs to
  rho_group1 <- rbernoulli(N_PARAMS, probs)
  rho_group2 <- !rho_group1

  # Keep track of how many values are in each group, used later for resampling
  n_group1 <- sum(rho_group1)
  n_group2 <- sum(rho_group2)

  # Resample values of rho that are larger than 1
  group1 <- rnorm(n_group1, mean = 1, sd = 0.05)
  n_bad1 <- sum(group1 > 1)
  while(n_bad1 > 0) {
    group1[group1 > 1] <- rnorm(n_bad1, mean = 1, sd = 0.175)
    n_bad1 <- sum(group1 > 1)
  }

  # get standard deviation for regression
  std_beta1 <- (beta[rho_group2] - mean(beta[rho_group2])) / sd(beta[rho_group2])

  group2 <- 0.25 - 0.025 * std_beta1 - (rlnorm(n_group2, sdlog = 0.2) - 1)

  # Resample values of rho in group2 as needed
  which_bad2 <- group2 < .13
  while(sum(which_bad2) > 0) {
    group2[group2 < 0.13] <- 0.25 - 0.1 * std_beta1[which_bad2] - (rlnorm(sum(which_bad2), sdlog = 0.2) - 1)

    which_bad2 <- group2 < .13
  }

  rhos[rho_group1] <- group1
  rhos[rho_group2] <- group2

  # Return the values of rho
  rhos
}
# Sample rho, using already sampled beta1.
starting_params$rho <- sample_rho(starting_params$beta1)

sample_tau <- function(mu = 4.1, n_params = N_PARAMS) {
  # Function used to sample tau. Tau is approximately independent of the other
  # parameters.
  #
  # params:
  #   mu: mean of the truncated normal distribution
  # returns:
  #   vector of length 300, sampled values of tau from truncated normal distribution

  # Sample from normal distribution
  X <- rnorm(n_params, mean = mu, sd = (1.7-mu) / qnorm(0.001))

  # Get how many resamples are needed
  n_resamp <- sum(X > 5.9 | X < 1.7)

  # Keep resampling until all values fall within desired range
  while (n_resamp > 0) {
    X[X > 5.9 | X < 1.7] <- rnorm(n_resamp, mean = mu, sd = (1.7-mu) / qnorm(0.001))

    n_resamp <- sum(X > 5.9 | X < 1.7)
  }

  # Return values of tau
  X
}
# Use function to sample tau
starting_params$tau <- sample_tau()

# Here we note that there are additional parameters that were fit, but their
# relationship between the remaining parameters is not given in Lee et al. (2020).
# Because of this, we simply sample these parameters using a truncated normal distribution:

POP <- 10911819
E_0_probs <- dnorm(seq(1, 1752, 1), mean = 400, sd = (1752-400) / qnorm(0.99))
I_0_probs <- dnorm(seq(1, 2271, 1), mean = 500, sd = (2271-500) / qnorm(0.99))
starting_params$E_0 <- sample(1:1752, size = N_PARAMS, replace = TRUE, prob = E_0_probs) / POP
starting_params$I_0 <- sample(1:2271, size = N_PARAMS, replace = TRUE, prob = I_0_probs) / POP
starting_params$S_0 <- 1 - starting_params$E_0 - starting_params$I_0

rm(sample_beta_nu, sample_rho, sample_tau, POP, I_0_probs, E_0_probs)
gc()

@@

bounds <- tribble(
  ~param, ~lower, ~upper,
  "beta2", 1e-08, 10,
  "beta3", 1e-08, 10,
  "beta4", 1e-08, 10,
  "beta5", 1e-08, 10,
  "beta6", 1e-08, 10
)

@@

betas <- sobol_design(
  lower = lower,
  upper = upper,
  nseq  = N_PARAMS
)

epi_params <- cbind(starting_params, betas)

not_in_start_params <- names(coef(h1_epi))[!names(coef(h1_epi)) %in% colnames(epi_params)]

for (i in 1:length(not_in_start_params)) {
  epi_params[, not_in_start_params[i]] <- coef(h1_epi)[not_in_start_params[i]]
}

rm(i)

@@

# Check if RUN_LEVEL == 3 results exists, if so, use them.
# Include a flag to check if RUN_LEVEL == 3. If so, we might want to recompute.
# TODO: Only save parameters and likelihoods, not entire MIF objects.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_fit.rds")) {
  lee1_epi_mif <- readRDS("model1/run_level_3/lee1_epi_fit.rds")
} else {
  lee1_epi_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_fit.rds"),
    {
      lee1_epi_mif_results <- list()

      foreach(
        i = 1:nrow(epi_params),
        .packages = c('pomp'),
        .combine = c
      ) %dopar% {
        r_params <- unlist(epi_params[i, ])
        coef(h1_epi) <- r_params
        mif2(
          h1_epi,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.fraction.50 = 0.5,  # Cooling set so that we will reach smaller rw.sd at the end than the end of the unit3 search.
          rw.sd = epi_rw
        )
      } -> lee1_epi_mif_out

      foreach(i = 1:length(lee1_epi_mif_out), .combine = rbind) %do% {
        guess <- coef(lee1_epi_mif_out[[i]])
        sims <- simulate(
          h1_epi,
          n = N_LEE_SIMS,
          format = 'data.frame',
          params = guess
        )  %>%
          dplyr::filter(week == max(week)) %>%
          dplyr::select(-.id, -week)

@@

rm(epi_rw, epi_params, not_in_start_params,
   lower, upper, betas, bounds, starting_params)
gc()

@@

\begin{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Informing policy via dynamic models: Eliminating cholera in Haiti}
\runtitle{Eliminating Cholera in Haiti}

\begin{aug}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only one address is permitted per author. %%
%% Only division, organization and e-mail is %%
%% included in the address.                  %%
%% Additional information can be included in %%
%% the Acknowledgments section if necessary. %%
%% ORCID can be inserted by command:         %%
%% \orcid{0000-0000-0000-0000}               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[A]{\fnms{Jesse}~\snm{Wheeler}\ead[label=e1]{jeswheel@umich.edu}\orcid{0000-0003-3941-3884}},
\author[A]{\fnms{AnnaElaine}~\snm{Rosengart}\ead[label=e2]{aelr@umich.edu}}
\author[A]{\fnms{Zhuoxun}~\snm{Jiang}\ead[label=e3]{zhuoxunj@umich.edu}},
\author[A]{\fnms{Kevin}~\snm{Hao En Tan}\ead[label=e4]{kevtan@umich.edu}},
\author[A]{\fnms{Noah}~\snm{Treutle}\ead[label=e5]{ntreutle@umich.edu}}
\and
\author[A]{\fnms{Edward}~\snm{Ionides}\ead[label=e6]{ionides@umich.edu}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Addresses                                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address[A]{Statistics Department, University of Michigan\printead[presep={,\ }]{e1,e2,e3,e4,e5,e6}}
\end{aug}

@@

\begin{figure}[ht]
<<Plot_Reported_Cases, echo=FALSE, fig.height=3.5>>=
plot_df <- haitiCholera %>%
  select(-report) %>%
  mutate(date = as.Date(date_saturday)) %>%
  select(-date_saturday) %>%
  pivot_longer(
    data = .,
    cols = -c(date),
    names_to = 'Departement',
    values_to = "Cases",
  )

ggplot(plot_df, aes(x = date, y = Cases + 1)) +
  facet_wrap(~Departement, nrow = 2, labeller = dep_labeller) +
  geom_line() +
  theme(
    axis.title.x = element_blank()
  ) +
  ylab('Reported Cases') +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2011-01-01"), as.Date("2019-01-01"), by = '2 years'))

@
\caption{\label{CholeraData}
\small
Weekly reported cholera cases in Haiti from October 2010 to January 2019.
}
\end{figure}

@@

<<FitModel1, echo=FALSE, message=FALSE, include=FALSE>>=
set.seed(636813)

NP_H1        <- switch(RUN_LEVEL, 100, 1000,   2000)
NMIF_H1      <- switch(RUN_LEVEL,  10,   50,    200)
NUM_TREND    <- switch(RUN_LEVEL,   5,   10,     90)
NPROF        <- switch(RUN_LEVEL,   4,    4,     24)
NP_EVAL      <- switch(RUN_LEVEL, 200, 1000,   5000)
NREPS_EVAL   <- switch(RUN_LEVEL, cores, 20,  cores)

if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/haiti1_fit.rds")) {
  h1_MIF <- readRDS("model1/run_level_3/haiti1_fit.rds")
} else {
  h1_MIF <- bake(
    file = paste0("model1/", rl_dir, "haiti1_fit.rds"), {
      fit_haiti1(
        NP = NP_H1,
        NMIF = NMIF_H1,
        NUM_TREND = NUM_TREND,
        NPROF = NPROF,
        NREPS_EVAL = NREPS_EVAL,
        NP_EVAL = NP_EVAL,
        ncores = cores
      )
    }
  )
}

h1_prof_res <- h1_MIF %>%
  group_by(betat) %>%
  summarize(logLik = max(ll))

@@

mod1_ll <- h1_MIF %>%
  filter(ll == max(ll)) %>%
  select(ll) %>%
  unlist()

rm(NP_H1, NMIF_H1, NUM_TREND, NPROF, NREPS_EVAL, NP_EVAL, h1_MIF)
gc()

@@

<<FitModel3, echo=FALSE, include=FALSE, message=FALSE>>=
set.seed(1859687)

@@

rm(
  SEARCH1, SEARCH2, SEARCH3, N_SEARCHES, unit_specific_names, 
  shared_param_names, est_param_names, est_param_names_expanded, 
  setrw1, setrw2, setrw3, reg_rw_1.sd, reg_rw_2.sd, reg_rw_3.sd, 
  chol_rw_1.sd, chol_rw_2.sd, chol_rw_3.sd
)
gc()

<<Load Model Parameters, echo=FALSE, results='hide'>>=

@@

# Fit model 2. Note that this takes the same amount of time at all run
# levels.
h2_fit <- bake(
  file = paste0("model2/model2_fit.rds"), {
    fit_haiti2()
  },
  timing = FALSE
)

@@

best_m3 <- which.max(haiti3_fit[[length(haiti3_fit)]]$logLiks$logLik)

p3 <- haiti3_fit[[length(haiti3_fit)]]$params[best_m3, ]

@@

\begin{figure}[ht]
\centering
<<Model_1_Sims_Figure, fig.height=2.4, fig.width=4.7, fig.align='center'>>=
sims <- simulate(h1, nsim = 500, format = 'data.frame', seed = 3448931)

quants <- sims %>%
  mutate(is_data = .id == 'data') %>%
  filter(!is_data) %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

ggplot() +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = quants, aes(x = date, ymin = q025 + 1, ymax = q975 + 1), alpha = 0.5) +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 8),
        axis.title.y = element_text(size = 10))+
  ylab('Reported cholera cases') +
  geom_vline(xintercept = lubridate::weeks(232) + lubridate::ymd("2010-10-16"), linetype = 'dashed') +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2011-01-01"), as.Date("2019-01-01"), by = '1 year'))
@
\caption{\label{fig:mod1fit}
Simulations from Model~1 compared to reported cholera cases.
The black curve is observed data, the blue curve is median of 500 simulations from the fitted model, and the vertical dashed line represents break-point when parameters are refit.
}
\end{figure}

@@

<<Model 2 Trajectory, echo=FALSE>>=
h2_epi_traj <- trajectory(h2_epi, params = h2_fit$h2_params, format = 'data.frame')
h2_epi_traj$Ctotal <- rowSums(h2_epi_traj[, paste0("C", 1:10)]) * h2_fit$h2_params['Rho']

@@

<<Model 3 SpatPOMP sims, echo=FALSE>>=
h3_spat_nsim <- switch(RUN_LEVEL, 20, 100, 500)

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPompSims.rds")) {
  h3Spat_quants <- readRDS("model3/run_level_3/SpatPompSims.rds")
} else {
  h3Spat_quants <- bake(
    file = paste0('model3/', rl_dir, 'SpatPompSims.rds'), {
      # h3Spat <- haiti3_spatPomp()
      # coef(h3Spat) <- haitipkg:::panelParms_toSpatParms3(p3)

      h3Spat_sims <- simulate(
        h3Spat, nsim = h3_spat_nsim, format = 'data.frame'
      )

>>>>

<<Setup, include=FALSE,echo=FALSE,results='hide'>>=

# Load Required packages
library(knitr)
library(pomp)
library(panelPomp)
library(spatPomp)
library(foreach)
library(doParallel)
library(doRNG)
library(haitipkg)
library(tidyverse)

# Set run level:
#   1: Used primarily for debugging
#   2: Used to test various results in a way that provides meaningful results, yet doesn't require an extensive amount of time to run.
#   3: Used to get final results.
#
# Throughout the document, if run_level_3 results exist already exist,
# then these results are used.
RUN_LEVEL <- 1
rl_dir <- paste0("run_level_", RUN_LEVEL, "/")

# Check if run_level_* directories exist; if not, create them.
for (i in c(1, 3)) {
  if (!dir.exists(paste0("model", i, '/', rl_dir))) {
    dir.create(paste0("model", i, '/', rl_dir), recursive = TRUE)
  }
}

# Model 2 doesn't need a run_level, since all computations are the same.
if (!dir.exists('model2')) dir.create("model2/")

@@

# Setting black and white ggplot2 theme for entire document.
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5)))

# Get the number of cores. The high performance computing cluster that we
# used has 36 cores.
cores <-  as.numeric(Sys.getenv('SLURM_NTASKS_PER_NODE', unset=NA))
if(is.na(cores)) cores <- detectCores()
registerDoParallel(cores)

# Create a mapping between department names and what should be shown in figures
dep_labeller <- as_labeller(
  c(
    'Artibonite' = 'Artibonite',
    'Sud_Est' = 'Sud-Est',
    'Sud.Est' = 'Sud-Est',
    'Nippes' = 'Nippes',
    'Nord_Est' = 'Nord-Est',
    'Nord.Est' = 'Nord-Est',
    'Ouest' = 'Ouest',
    'Centre' = 'Centre',
    'Nord' = 'Nord',
    'Sud' = 'Sud',
    'Nord_Ouest' = 'Nord-Ouest',
    'Nord.Ouest' = 'Nord-Ouest',
    'Grande_Anse' = 'Grand\'Anse',
    'Grand.Anse' = 'Grand\'Anse'
  )
)

# Vector used to arrange figures in the same order that was used in Lee et al (2020)
plot_order <- c(
  'Artibonite',
  'Sud_Est',
  'Nippes',
  'Nord_Est',
  'Ouest',
  'Centre',
  'Nord',
  'Sud',
  'Nord_Ouest',
  'Grande_Anse'
)

@@

# Create "long" version of the observed cholera cases, for plotting purposes.
dep_plot_df <- haitiCholera %>%
  select(-report) %>%
  pivot_longer(
    data = .,
    cols = -date_saturday,
    values_to = 'cases',
    names_to = 'dep'
  ) %>%
  mutate(
    date = as.Date(date_saturday),
    dep = gsub("\\.", "_", dep)
  ) %>%
  mutate(
    dep = case_when(dep == "Grand_Anse" ~ "Grande_Anse", TRUE ~ dep)
  )

# Calculate total number of nationally aggregated cholera cases, using previous
# data.
true_agg_cases <- dep_plot_df %>%
  tidyr::pivot_wider(
    data = .,
    id_cols = c(date),
    names_from = dep,
    values_from = cases,
    names_prefix = 'cases_'
  ) %>%
  mutate(
    ReportedAll = cases_Artibonite + cases_Centre +
      cases_Grande_Anse + cases_Nippes + cases_Nord +
      cases_Nord_Est + cases_Ouest + cases_Sud +
      cases_Sud_Est + cases_Nord_Ouest
  )
@

@@

set.seed(141837)

# Get starting parameters close to those that are provided
# in Lee et al (2020)
starting_params <- get_h1_startVals(N_PARAMS)

@@

# Do a small global search for starting parameters for beta2-beta6
bounds <- tribble(
  ~param, ~lower, ~upper,
  "beta2", 1e-08, 10,
  "beta3", 1e-08, 10,
  "beta4", 1e-08, 10,
  "beta5", 1e-08, 10,
  "beta6", 1e-08, 10
)

@@

# Get a sequence of starting values for the betas
betas <- sobol_design(
  lower = lower,
  upper = upper,
  nseq  = N_PARAMS
)

# Combine betas with the starting_params data.frame object.
epi_params <- cbind(starting_params, betas)

# Get the names of all of the remaining coefficients (fixed parameters).
not_in_start_params <- names(coef(h1_epi))[!names(coef(h1_epi)) %in% colnames(epi_params)]

# Add fixed parameters to the matrix of starting parameters
for (i in 1:length(not_in_start_params)) {
  epi_params[, not_in_start_params[i]] <- coef(h1_epi)[not_in_start_params[i]]
}

@@

# Check if RUN_LEVEL == 3 results exists, if so, use them.
# Include a flag to check if RUN_LEVEL == 3. If so, we might want to recompute.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_fit.rds")) {
  lee1_epi_mif <- readRDS("model1/run_level_3/lee1_epi_fit.rds")
} else {
  lee1_epi_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_fit.rds"),
    {
      # Create a list to save all of the results.
      lee1_epi_mif_results <- list()

      # Loop through each of the starting parameters
      foreach(
        i = 1:nrow(epi_params),
        .packages = c('pomp'),
        .combine = c
      ) %dopar% {

        # Get the ith set of starting parameters
        r_params <- unlist(epi_params[i, ])

        # Set the coefficients of model 1 to the ith starting values
        coef(h1_epi) <- r_params

        # Perform short IF2 from the starting value
        mif2(
          h1_epi,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.fraction.50 = 0.5,
          rw.sd = epi_rw
        )
      } -> lee1_epi_mif_out  # Contains nrow(epi_params) mif2d objects

      # For each output of the MIF2 objects, simulate Model 1 with the IF2
      # parameters and save the states from the final week of the simulation
      foreach(i = 1:length(lee1_epi_mif_out), .combine = rbind) %do% {

        # Get IF2 parameter set
        guess <- coef(lee1_epi_mif_out[[i]])

        # Simulate with the parameter set
        sims <- simulate(
          h1_epi,
          n = N_LEE_SIMS,
          format = 'data.frame',
          params = guess
        )  %>%
          dplyr::filter(week == max(week)) %>%  # Just keep final states, to use later.
          dplyr::select(-.id, -week)

@@

rm(epi_rw, epi_params, not_in_start_params,
   lower, upper, betas, bounds, starting_params, i)
gc()

@@

\begin{frontmatter}

\title{Informing policy via dynamic models: Eliminating cholera in Haiti}
\runtitle{Eliminating Cholera in Haiti}

\begin{aug}

\author[A]{\fnms{Jesse}~\snm{Wheeler}\ead[label=e1]{jeswheel@umich.edu}\orcid{0000-0003-3941-3884}},
\author[A]{\fnms{AnnaElaine}~\snm{Rosengart}\ead[label=e2]{aelr@umich.edu}}
\author[A]{\fnms{Zhuoxun}~\snm{Jiang}\ead[label=e3]{zhuoxunj@umich.edu}},
\author[A]{\fnms{Kevin}~\snm{Hao En Tan}\ead[label=e4]{kevtan@umich.edu}},
\author[A]{\fnms{Noah}~\snm{Treutle}\ead[label=e5]{ntreutle@umich.edu}}
\and
\author[A]{\fnms{Edward}~\snm{Ionides}\ead[label=e6]{ionides@umich.edu}}

\address[A]{Statistics Department, University of Michigan\printead[presep={,\ }]{e1,e2,e3,e4,e5,e6}}

\end{aug}

@@

\begin{figure}[ht]
<<Plot_Reported_Cases, echo=FALSE, fig.height=3.5>>=

# Plot reported cholera cases, by department. Creates Figure 1.
ggplot(dep_plot_df, aes(x = date, y = cases + 1)) +
  facet_wrap(~dep, nrow = 2, labeller = dep_labeller) +
  geom_line() +
  theme(axis.title.x = element_blank()) +
  ylab('Reported Cases') +
  scale_y_log10(
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  )
@
\caption{\label{CholeraData}
\small
Weekly reported cholera cases in Haiti from October 2010 to January 2019.
}
\end{figure}

@@

<<FitModel1, echo=FALSE, message=FALSE, include=FALSE>>=

# This code chunk is used to fit Model 1 using the IF2 algorithm.
set.seed(636813)

# Check RUN_LEVEL, and set hyper-parameters to IF2 and pfilter accordingly.
#
#                      RUN_LEVEL =    1,    2,      3
NP_H1        <- switch(RUN_LEVEL,   100, 1000,   2000)
NMIF_H1      <- switch(RUN_LEVEL,    10,   50,    200)
NUM_TREND    <- switch(RUN_LEVEL,     5,   10,     90)
NPROF        <- switch(RUN_LEVEL,     4,    4,     24)
NP_EVAL      <- switch(RUN_LEVEL,   200, 1000,   5000)
NREPS_EVAL   <- switch(RUN_LEVEL, cores,   20,  cores)

# Check first if RUN_LEVEL_3 file exists in directory, if not, create the
# RUN_LEVEL version of the results and save them.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/haiti1_fit.rds")) {
  h1_MIF <- readRDS("model1/run_level_3/haiti1_fit.rds")
} else {
  h1_MIF <- bake(
    file = paste0("model1/", rl_dir, "haiti1_fit.rds"), {
      fit_haiti1(
        NP = NP_H1,
        NMIF = NMIF_H1,
        NUM_TREND = NUM_TREND,
        NPROF = NPROF,
        NREPS_EVAL = NREPS_EVAL,
        NP_EVAL = NP_EVAL,
        ncores = cores
      )
    }
  )
}

# The result is a data.frame containing many parameter values and their
# corresponding likelihood evaluation. To perform a likelihood profile,
# we calculate the maximum likelihood for each value of betat.
h1_prof_res <- h1_MIF %>%
  group_by(betat) %>%
  summarize(logLik = max(ll))

@@

# Save the likelihood of the best parameters, for use later.
mod1_ll <- h1_MIF %>%
  filter(ll == max(ll)) %>%
  select(ll) %>%
  unlist()

# Just a bit of tidying up: removing parameters that we won't use again.
rm(NP_H1, NMIF_H1, NUM_TREND, NPROF, NREPS_EVAL, NP_EVAL, h1_MIF)
gc()

@@

<<FitModel2, echo=FALSE, include=FALSE, message=FALSE>>=

# This code chunk fits Model 2. This is a quick caclulation compared to models
# 1 and 3, so no run_level is used.

h2_fit <- bake(
  file = paste0("model2/model2_fit.rds"), {
    fit_haiti2()
  },
  timing = FALSE
)
@

<<FitModel3, echo=FALSE, include=FALSE, message=FALSE>>=

# This code chunk is used to fit Model 3 using the IBPF algorithm.
set.seed(1859687)

@@

@@

# Just a bit of tidying up: removing parameters that we won't use again.
rm(
  SEARCH1, SEARCH2, SEARCH3, N_SEARCHES, unit_specific_names,
  shared_param_names, est_param_names, est_param_names_expanded,
  setrw1, setrw2, setrw3, reg_rw_1.sd, reg_rw_2.sd, reg_rw_3.sd,
  chol_rw_1.sd, chol_rw_2.sd, chol_rw_3.sd
)
gc()

@@

# Find the index for parameters of model 3 that correspond to the MLE.
best_m3 <- which.max(haiti3_fit[[length(haiti3_fit)]]$logLiks$logLik)

# Save the parameters that correspond to the MLE of Model 3.
p3 <- haiti3_fit[[length(haiti3_fit)]]$params[best_m3, ]

@@

\begin{figure}[ht]
\centering
<<Model_1_Sims_Figure, fig.height=2.4, fig.width=4.7, fig.align='center'>>=
# This code chunk creates the figure containing simulations from Model 1

# Simulate Model 1 from initial conditions.
sims <- simulate(h1, nsim = 500, format = 'data.frame', seed = 3448931)

# Calculate 0.025 and 0.975 percentiles of simulations, at each time point.
quants <- sims %>%
  mutate(is_data = .id == 'data') %>%
  filter(!is_data) %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

# Plot the data, median value of the simulation, and estimated quantiles.
ggplot() +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(  # Plot confidence region
    data = quants,
    aes(x = date, ymin = q025 + 1, ymax = q975 + 1),
    alpha = 0.5
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 8),
    axis.title.y = element_text(size = 10)
  )+
  ylab('Reported cholera cases') +
  geom_vline(  # Show change from epidemic -> endemic models.
    xintercept = lubridate::weeks(232) + lubridate::ymd("2010-10-16"),
    linetype = 'dashed'
  ) +
  scale_y_log10(  # Plot on the log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(  # Format the date on x-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      as.Date("2019-01-01"),
      by = '1 year'
    )
  )
@
\caption{\label{fig:mod1fit}
Simulations from Model~1 compared to reported cholera cases.
The black curve is observed data, the blue curve is median of 500 simulations from the fitted model, and the vertical dashed line represents break-point when parameters are refit.
}
\end{figure}

@@

<<Model 2 Trajectory, echo=FALSE>>=

# This code chunk calculates the "trajectory" of the differential equations
# for Model 2.

h2_epi_traj <- trajectory(
  h2_epi,  # Model
  params = h2_fit$h2_params,  # estimated parameters
  format = 'data.frame'
)

# Display the output as nationally aggregated data, as was done in Lee et al (2020)
h2_epi_traj$Ctotal <- rowSums(
  h2_epi_traj[, paste0("C", 1:10)]  # Infections in each department
) * h2_fit$h2_params['Rho']  # multiply by reporting rate, to compare to reported cases.

@@

<<Model 3 SpatPOMP sims, echo=FALSE>>=

# Number of simulations from Model 3. Because it's high-dimensional, we include
# a run level to avoid time/memory issues when editing. Final results shown
# with RUN_LEVEL = 3.

h3_spat_nsim <- switch(RUN_LEVEL, 20, 100, 500)

# Check if run_level_3 results exit, and use them if they do.
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPompSims.rds")) {
  h3Spat_quants <- readRDS("model3/run_level_3/SpatPompSims.rds")
} else {
  h3Spat_quants <- bake(
    file = paste0('model3/', rl_dir, 'SpatPompSims.rds'), {

      # Simulate from the fitted version of Model 3
      h3Spat_sims <- simulate(
        h3Spat, nsim = h3_spat_nsim, format = 'data.frame'
      )
